---
title: "More interactive mapping with R"
subtitle: "<br>shorturl.at/jCX09<br>"
author: "Duc-Quang Nguyen"
date: "2022/05/19"
output: 
  revealjs::revealjs_presentation:
    theme: serif
    echo: TRUE
    warning: FALSE
    message: FALSE
---



```{r load packages,warning=FALSE,message=FALSE}
library(tidyverse)
library(sf)
library(swissdd)
library(leaflet)
library(htmltools)
#install.packages(c("htmltools", leaflet", "highcharter"))
```

# Leaflet

Leaflet is one of the most popular open-source JavaScript libraries for interactive maps

+  Good performance, many options, very large support/users, maintained by RStudio folks
-  API syntax not always straightforward to learn from R 

* [Leaflet](https://rstudio.github.io/leaflet/)
* [Basemaps](http://leaflet-extras.github.io/leaflet-providers/preview/index.html)

---

## Choropleth map by municipality of the federal ballot «Lex Netflix» 
 (again)


```{r load votation data and swiss municipalities boundaries, warning=FALSE,message=FALSE}
muni <- st_read("shp/g2g22.shp")

ln_df  <- swissdd::get_nationalvotes(
  geolevel = "municipality", from_date = "2022-05-15", 
  to_date =  "2022-05-15", language = "FR") %>% 
  filter(id == 6550) %>% 
  select(canton_name:mun_name, jaStimmenInProzent, gueltigeStimmen)

ln_df <- left_join(
  ln_df %>% mutate(mun_id = as.numeric(mun_id)), 
  muni %>% select(GMDNR:GRNR), 
  by = c("mun_id" = "GMDNR")
  # To ensure the join returns a sf object
) %>% st_as_sf()

```

## Reproject the data

```{r reproject and tooltip}
# reproject into WGS84 coordinates: https://rspatial.org/raster/spatial/6-crs.html
ln_df <- ln_df %>% 
  st_transform(crs = 4326) %>% 
   mutate(
    tp = str_c("<h3>", mun_name, "<h3>",
               "<i>", canton_name, "</i><br>",
               "<b>",
               prettyNum(jaStimmenInProzent, 
                         decimal.mark = ",", digits = 3),
               "% oui</b>"
               ) %>% lapply(htmltools::HTML) # important somehow...
   )
```


## Lealet map

```{r leaflet map}
# Function to discretize and colour % yes vote
pal <- colorBin(
  "PRGn", domain = ln_df$jaStimmenInProzent , 
  bins = seq(0, 100, 10))

lmap <- leaflet(ln_df) %>%
  # Define the basemap 
  addProviderTiles(providers$CartoDB.PositronNoLabels,
                   options = providerTileOptions(opacity = 1)) %>%
  addPolygons( 
    color = "white", weight = 0.2, smoothFactor = 0.5,
    opacity = 0.95, fillOpacity = 0.98,
    fillColor = ~pal(jaStimmenInProzent),
    label = ln_df$tp
  ) %>% 
  # Add basemap labels layer
  addMapPane("labels", zIndex = 410) %>% # show labels above
  addProviderTiles(providers$CartoDB.PositronOnlyLabels,
                   options = providerTileOptions(pane = "labels") ) %>%
  # Legend box
  addLegend(
    pal = pal, values = ~jaStimmenInProzent, opacity = 0.7,
    position = "bottomleft",
    title = "% de oui à Lex Netflix") %>%
  # Localized autozoom 
  addEasyButton(easyButton(
    icon="fa-crosshairs", title="Locate Me",
    onClick=JS("function(btn, map){ map.locate({setView: true}); }")))


```

---

```{r view leaflet map}

lmap

```

---

```{r add cantonal borders layer, results='hide'}
cant <- st_read("shp/g2k22.shp") %>% 
      st_transform(crs = 4326)

lmap %>% 
  addPolylines(
    data = cant,
    color = "grey", weight = 1.6, smoothFactor = 0.3,
    opacity = 0.7
  ) 

```



# Exercise 

* Load the cantonal tilemap ("shp/Switzerland_Tiles_CH1903LV03 ... ") [tilemap for Switzerland](https://github.com/ebp-group/Switzerland_Tilemap)

* Overlay it over a leaflet basemap. Bonus point if you chose another dark basemap ([Basemaps](http://leaflet-extras.github.io/leaflet-providers/preview/index.htm)



